---
title: "The Olympic Halo: Spatial Spillovers on Mountain Real Estate"
author: "Pietro Fraccaroli"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 7
)
```

```{r libraries}
library(tidyverse)
library(sf)
library(xml2)
library(ggplot2)
library(viridis)
library(scales)
library(fixest)
library(patchwork)
library(here)
library(ggspatial)
```

---

# 1. Load and Harmonise All OMI Price Data

```{r load-prices}
raw_dir <- here::here("Raw")

# Parse year and semester from filename: e.g. QIP_1380682_1_20251_VALORI.csv -> 2025, 1
parse_semester <- function(filepath) {
  fname <- basename(filepath)
  m     <- regmatches(fname, regexpr("_(\\d{4})(\\d)_VALORI", fname))
  year  <- as.integer(substr(m, 2, 5))
  sem   <- as.integer(substr(m, 6, 6))
  list(year = year, sem = sem)
}

valori_files <- list.files(raw_dir, pattern = "_VALORI\\.csv$",
                           recursive = TRUE, full.names = TRUE)

load_valori <- function(f) {
  ps <- parse_semester(f)
  # Row 1 is a title line; row 2 is the header
  df <- read.csv2(f, skip = 1, header = TRUE,
                  stringsAsFactors = FALSE, encoding = "UTF-8",
                  na.strings = c("", "NA", "n.d."))
  # Drop trailing empty columns produced by trailing semicolons
  df <- df[, colSums(!is.na(df)) > 0]
  df$year <- ps$year
  df$sem  <- ps$sem
  df
}

prices_raw <- do.call(rbind, lapply(valori_files, load_valori))

cat("Total rows loaded:", nrow(prices_raw), "\n")
cat("Semesters:", paste(sort(unique(paste0(prices_raw$year, "/", prices_raw$sem))), collapse = " "), "\n")
```
Perché ~44k righe?
Ogni riga è una combinazione unica di: zona OMI × tipologia immobile × semestre. Belluno ha ~60 comuni, ognuno con 2-6 zone OMI, e per ogni zona ci sono ~8-10 tipologie (abitazioni civili, box, negozi, uffici, capannoni...).

```{r}
head(prices_raw)
```
Colonna	Cosa indica
Area_territoriale	Macro-area geografica (sempre NORD-EST)
Regione	            Regione (sempre VENETO)
Prov	            Sigla provincia (sempre BL = Belluno)
Comune_ISTAT	    Codice ISTAT del comune (es. 5025006 = Belluno)
Comune_cat	        Codice catastale del comune (es. D2AA)
Comune_amm	        Codice belfiore del comune (es. A757 = Belluno, A266 = Cortina) — è la chiave che useremo per il join con i KML
Comune_descrizione	Nome del comune in chiaro
Fascia	            Fascia di mercato: B=centrale, D=semicentrale, E=periferica/extraurbana, R=rurale
Zona	            Codice zona OMI all'interno del comune (es. B1, B2, D3...)
LinkZona	        Codice univoco nazionale della zona (es. BL00000030)
Cod_Tip	            Codice tipologia immobile (20=abitazioni civili, 21=economiche, 13=box, 5=negozi...)
Descr_Tipologia  	Descrizione testuale della tipologia
Stato	            Stato conservativo: NORMALE, OTTIMO, SCADENTE
Stato_prev	        Stato prevalente nella zona (P = prevalente)
Compr_min / Compr_max	Prezzi minimi e massimi di acquisto in €/m²
Sup_NL_compr	    Superficie di riferimento per compravendita (L = lorda)
Loc_min / Loc_max	Canoni minimi e massimi di locazione in €/m²/mese
Sup_NL_loc	        Superficie di riferimento per locazione
year / sem	        Anno e semestre aggiunti da noi
```{r clean-prices}
prices <- prices_raw %>%
  # Residential civil (Cod_Tip == 20), normal condition
  filter(Cod_Tip == 20, Stato == "NORMALE") %>%
  mutate(
    price_mid = (Compr_min + Compr_max) / 2,           # midpoint purchase price €/m2
    zone_key  = paste0(Comune_amm, "_", Zona),          # composite key matching KML
    time      = year + (sem - 1) * 0.5,                 # numeric semester (e.g. 2019.5 = 2019/S2)
    post      = as.integer(time >= 2019.5)              # Olympic bid announced June 2019
  ) %>%
  filter(!is.na(price_mid)) %>%
  select(year, sem, time, post,
         Comune_ISTAT, Comune_amm, Comune_descrizione,
         Fascia, Zona, LinkZona, zone_key,
         price_mid, Compr_min, Compr_max)

cat("Observations after filter:", nrow(prices), "\n")
cat("Municipalities:", n_distinct(prices$Comune_amm), "\n")
cat("Unique zones:  ", n_distinct(prices$zone_key), "\n")
```

price_mid
I dati OMI non danno un prezzo esatto ma un intervallo. Es. Belluno zona B1 vale tra 1050 e 1350 €/m². Noi prendiamo il punto medio: (1050+1350)/2 = 1200 €/m². È la nostra variabile dipendente.

zone_key
Serve per fare il join tra i prezzi e le geometrie KML. Incolla il codice comune con il codice zona:

Comune_amm = A266, Zona = B1 → zone_key = "A266_B1"
Così ogni riga dei prezzi può essere collegata al poligono geografico corrispondente nel KML
time
Converte anno+semestre in un numero continuo per i grafici:

2019/S1 → 2019 + (1-1)×0.5 = 2019.0
2019/S2 → 2019 + (2-1)×0.5 = 2019.5
2020/S1 → 2020.0
2025/S2 → 2025.5
post
È la variabile di trattamento del DiD. Le Olimpiadi sono state assegnate a Milano-Cortina il 24 giugno 2019, quindi il primo semestre "trattato" è il 2019/S2 (luglio-dicembre 2019):

Prima del 2019/S2 → post = 0
Dal 2019/S2 in poi → post = 1

---

# 2. Build Spatial Layer from KML Files

Zone polygons come from the most recent OMI shapefile archive (BL20252.zip).
OMI zone boundaries change infrequently; this geometry serves as the stable
spatial reference for the full panel.

```{r read-kml-sf}
# Disable S2 engine globally before any spatial operation on these KML files.
# Some OMI polygons have unclosed rings that S2 rejects but GEOS tolerates.
sf_use_s2(FALSE)

zip_path <- file.path(raw_dir,
                      "QIP1380683_FRCPTR00T21F861Y",
                      "BL20252.zip")

tmp_dir <- file.path(tempdir(), "omi_kml")
dir.create(tmp_dir, showWarnings = FALSE)
unzip(zip_path, exdir = tmp_dir, overwrite = TRUE)

kml_files <- list.files(tmp_dir, pattern = "\\.kml$", full.names = TRUE)
cat("KML files extracted:", length(kml_files), "\n")

read_kml_sf <- function(f) {
  tryCatch({
    lyr    <- st_layers(f)$name[1]
    sf_obj <- st_read(f, layer = lyr, quiet = TRUE)
    if (nrow(sf_obj) == 0) return(NULL)
    sf_obj %>%
      mutate(kml_file = basename(f)) %>%
      select(Name, kml_file, geometry)
  }, error = function(e) NULL)
}

zones_sf_raw <- map(kml_files, read_kml_sf) %>%
  compact() %>%
  bind_rows()

cat("Zone features read:", nrow(zones_sf_raw), "\n")
```

```{r fix-broken-geometries}
# These 6 municipalities have broken geometries in BL20252.zip.
# We replace their KML features with the equivalent from BL20242.zip.
broken_codcom <- c("A206", "A982", "B375", "C672", "G404", "G642", "I876")

zip_fallback <- file.path(raw_dir,
                          "QIP1380681_FRCPTR00T21F861Y",
                          "BL20242.zip")

tmp_dir2 <- file.path(tempdir(), "omi_kml_fallback")
dir.create(tmp_dir2, showWarnings = FALSE)
unzip(zip_fallback, exdir = tmp_dir2, overwrite = TRUE)

kml_fallback <- list.files(tmp_dir2, pattern = "\\.kml$", full.names = TRUE)
kml_fallback <- kml_fallback[
  tools::file_path_sans_ext(basename(kml_fallback)) %in% broken_codcom
]

zones_fallback <- map(kml_fallback, read_kml_sf) %>%
  compact() %>%
  bind_rows()

cat("Fallback features read:", nrow(zones_fallback), "\n")

# Replace broken entries in zones_sf_raw with fallback geometries
zones_sf_raw <- zones_sf_raw %>%
  filter(!str_remove(kml_file, "\\.kml$") %in% broken_codcom) %>%
  bind_rows(zones_fallback)

cat("Zone features after fix:", nrow(zones_sf_raw), "\n")
```

```{r parse-zone-attributes}
# CODCOM comes from the KML filename (e.g. A266.kml -> A266)
# CODZONA is the last token after "Zona OMI " in the Placemark Name
zones_sf <- zones_sf_raw %>%
  mutate(
    CODCOM   = str_remove(kml_file, "\\.kml$"),
    CODZONA  = str_extract(Name, "(?<=Zona OMI )\\S+"),
    zone_key = paste0(CODCOM, "_", CODZONA)
  ) %>%
  filter(!is.na(CODZONA)) %>%
  st_zm(drop = TRUE) %>%
  st_make_valid() %>%
  st_transform(crs = 32632)   # UTM32N for metric distances

cat("Zones with valid attributes:", nrow(zones_sf), "\n")

# Centroids for distance calculation — with bounding box fallback for broken geometries
safe_centroid_coords <- function(geom) {
  # Try proper centroid first
  result <- tryCatch(
    st_coordinates(st_centroid(geom))[1, 1:2],
    error = function(e) NULL
  )
  if (!is.null(result)) return(result)
  # Fallback: midpoint of bounding box (works even on topologically broken geometries)
  tryCatch({
    bb <- st_bbox(geom)
    c(X = (bb["xmin"] + bb["xmax"]) / 2,
      Y = (bb["ymin"] + bb["ymax"]) / 2)
  }, error = function(e) c(NA_real_, NA_real_))
}

coords_list <- lapply(zones_sf$geometry, safe_centroid_coords)
coords_mat  <- do.call(rbind, coords_list)

zone_centroids <- zones_sf %>%
  st_drop_geometry() %>%
  mutate(cx = coords_mat[, 1], cy = coords_mat[, 2]) %>%
  filter(!is.na(cx)) %>%
  select(zone_key, CODCOM, CODZONA, cx, cy)

cat("Centroids computed:", nrow(zone_centroids), "\n")
```

---

# 3. Distance to Cortina d'Ampezzo

Cortina's OMI municipality code is **A266**.
We define the origin as the average centroid of all Cortina OMI zones.

```{r cortina-distance}
cortina_centroid <- zone_centroids %>%
  filter(CODCOM == "A266") %>%
  summarise(cx = mean(cx), cy = mean(cy))

cat("Cortina centroid (UTM32N):", round(cortina_centroid$cx), round(cortina_centroid$cy), "\n")

zone_centroids <- zone_centroids %>%
  mutate(
    dist_cortina_km = sqrt(
      (cx - cortina_centroid$cx)^2 + (cy - cortina_centroid$cy)^2
    ) / 1000
  )

summary(zone_centroids$dist_cortina_km)
```

---

# 4. Build the Analysis Panel

```{r build-panel}
panel <- prices %>%
  left_join(zone_centroids, by = "zone_key") %>%
  filter(!is.na(dist_cortina_km)) %>%
  mutate(
    dist_bin = cut(dist_cortina_km,
                   breaks = c(0, 20, 40, 60, 80, Inf),
                   labels = c("0-20 km", "20-40 km", "40-60 km",
                              "60-80 km", ">80 km"),
                   right  = FALSE),
    log_price = log(price_mid)
  )

cat("Panel rows:", nrow(panel), "\n")
cat("Zones matched:", n_distinct(panel$zone_key), "\n")


```

```{r}
View(panel)
```
```{r}
# Save to Processed folder
dir.create(here::here("Processed"), showWarnings = FALSE)
write.csv(panel, here::here("Processed", "panel.csv"), row.names = FALSE)
write.csv(zone_centroids, here::here("Processed", "zone_centroids.csv"), row.names = FALSE)
st_write(zones_sf %>% filter(st_is_valid(geometry)),
         here::here("Processed", "zones_sf.gpkg"),
         delete_dsn = TRUE, quiet = TRUE)
cat("Saved to Processed/panel.csv, zone_centroids.csv, zones_sf.gpkg\n")
```
---

# 5. Maps

## 5a. Distance Gradient across Belluno Province

```{r}
zones_map <- zones_sf %>%
  filter(st_is_valid(geometry)) %>%
  st_zm(drop = TRUE) %>%
  left_join(zone_centroids %>% select(zone_key, dist_cortina_km), by = "zone_key")

province_outline <- st_union(zones_map)

```
```{r map-setup}
# Load all processed data from disk — no need to re-run spatial pipeline
panel          <- read.csv(here::here("Processed", "panel.csv"))
zone_centroids <- read.csv(here::here("Processed", "zone_centroids.csv"))
zones_sf       <- st_read(here::here("Processed", "zones_sf.gpkg"), quiet = TRUE)

zones_map <- zones_sf %>%
  left_join(zone_centroids %>% select(zone_key, dist_cortina_km),
            by = "zone_key")

province_outline <- st_union(zones_map)

cortina_pt <- st_sfc(st_point(c(
  zone_centroids %>% filter(CODCOM == "A266") %>% summarise(cx = mean(cx)) %>% pull(cx),
  zone_centroids %>% filter(CODCOM == "A266") %>% summarise(cy = mean(cy)) %>% pull(cy)
)), crs = 32632)
```

## 5a. Zoom: Cortina d'Ampezzo OMI Zones

```{r map-cortina-zoom, fig.height=8, fig.width=9}
cortina_zones <- zones_map %>%
  filter(grepl("^A266_", zone_key)) %>%
  mutate(zona = sub("A266_", "", zone_key))

cortina_outline <- st_union(cortina_zones)

centroids_coords <- st_coordinates(st_centroid(cortina_zones))
cortina_zone_labels <- cortina_zones
cortina_zone_labels$cx <- centroids_coords[, 1]
cortina_zone_labels$cy <- centroids_coords[, 2]

bbox <- st_bbox(cortina_zones)
pad  <- 3000  # 3 km padding around Cortina

ggplot() +
  geom_sf(data = cortina_zones,
          fill = "steelblue", alpha = 0.25,
          colour = "steelblue4", linewidth = 0.6) +
  geom_sf(data = cortina_outline,
          fill = NA, colour = "black", linewidth = 1.2) +
  geom_text(data = cortina_zone_labels,
            aes(x = cx, y = cy, label = zona),
            size = 4, fontface = "bold", colour = "black") +
  geom_sf(data = cortina_pt,
          colour = "red", size = 5, shape = 18) +
  coord_sf(
    xlim = c(bbox["xmin"] - pad, bbox["xmax"] + pad),
    ylim = c(bbox["ymin"] - pad, bbox["ymax"] + pad)
  ) +
  labs(
    title    = "Cortina d'Ampezzo — OMI Zones (A266)",
    subtitle = "Ogni poligono = una zona OMI | Contorno nero = confine comunale | Diamante rosso = centroide",
    caption  = "Source: OMI / Agenzia delle Entrate (2025)"
  ) +
  theme_void(base_size = 12) +
  theme(
    plot.title    = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 9, colour = "grey40")
  )
```

## 5a-bis. Cortina OMI Zones — Data Coverage Table

```{r cortina-coverage-table}
total_semesters <- n_distinct(panel$time)

cortina_summary <- panel %>%
  filter(CODCOM == "A266") %>%
  group_by(zone_key) %>%
  summarise(
    n_semesters   = n_distinct(time),
    first_period  = min(time),
    last_period   = max(time),
    mean_price    = round(mean(price_mid, na.rm = TRUE), 0),
    min_price     = round(min(price_mid, na.rm = TRUE), 0),
    max_price     = round(max(price_mid, na.rm = TRUE), 0),
    .groups = "drop"
  ) %>%
  mutate(
    zona        = sub("A266_", "", zone_key),
    coverage    = paste0(n_semesters, "/", total_semesters),
    price_range = paste0(min_price, " – ", max_price, " €/m²")
  ) %>%
  arrange(zona) %>%
  select(zona, coverage, first_period, last_period, mean_price, price_range)

# All A266 zones in geometry (including those with no prices)
all_cortina_zones <- zones_map %>%
  filter(grepl("^A266_", zone_key)) %>%
  mutate(zona = sub("A266_", "", zone_key)) %>%
  st_drop_geometry() %>%
  select(zone_key, zona)

coverage_table <- all_cortina_zones %>%
  left_join(cortina_summary, by = "zona") %>%
  mutate(
    coverage   = if_else(is.na(coverage), paste0("0/", total_semesters), coverage),
    mean_price = if_else(is.na(mean_price), NA_real_, mean_price),
    price_range = if_else(is.na(price_range), "— nessun dato —", price_range)
  ) %>%
  arrange(zona) %>%
  select(zona, coverage, first_period, last_period, mean_price, price_range)

knitr::kable(
  coverage_table,
  col.names = c("Zona OMI", "Semestri con dati", "Primo", "Ultimo",
                "Prezzo medio (€/m²)", "Range prezzi"),
  caption = "Cortina d'Ampezzo (A266) — Copertura dati OMI per zona"
)
```

## 5b. OMI Zones — No Fill

```{r map-zones-plain, fig.height=9}
ggplot() +
  geom_sf(data = zones_map,
          fill = "grey95", colour = "grey40", linewidth = 0.15) +
  geom_sf(data = province_outline,
          fill = NA, colour = "black", linewidth = 0.6) +
  geom_sf(data = cortina_pt,
          colour = "red", size = 4, shape = 18) +
  labs(
    title    = "Belluno Province: OMI Zones",
    subtitle = "Red diamond = Cortina d'Ampezzo",
    caption  = "Source: OMI / Agenzia delle Entrate (2025)"
  ) +
  theme_void() +
  theme(
    plot.title    = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, colour = "grey40")
  )
```

## 5b. Price Heat Map — Spatial Evolution Over Time

```{r map-prices-faceted, fig.height=9, fig.width=16}
# Log price for 4 anchor semesters
get_log_price <- function(yr, sm) {
  panel %>%
    filter(year == yr, sem == sm) %>%
    group_by(zone_key) %>%
    summarise(lp = mean(log(price_mid), na.rm = TRUE), .groups = "drop")
}

lp_2014S1 <- get_log_price(2014, 1)
lp_2019S1 <- get_log_price(2019, 1)
lp_2019S2 <- get_log_price(2019, 2)
lp_2025S1 <- get_log_price(2025, 1)

# Log growth = log(end) - log(start) ≈ % change
growth_pre <- lp_2014S1 %>%
  inner_join(lp_2019S1, by = "zone_key", suffix = c("_start", "_end")) %>%
  mutate(log_growth = lp_end - lp_start,
         label = "Pre-announcement\n(2014/S1 → 2019/S1)")

growth_post <- lp_2019S2 %>%
  inner_join(lp_2025S1, by = "zone_key", suffix = c("_start", "_end")) %>%
  mutate(log_growth = lp_end - lp_start,
         label = "Post-announcement\n(2019/S2 → 2025/S1)")

growth_all <- bind_rows(growth_pre, growth_post) %>%
  mutate(label = factor(label, levels = c(
    "Pre-announcement\n(2014/S1 → 2019/S1)",
    "Post-announcement\n(2019/S2 → 2025/S1)"
  )))

# Symmetric color scale capped at ±15% (log ≈ 0.15); outliers squished to boundary color
scale_cap <- 0.05

zones_growth <- do.call(rbind, lapply(levels(growth_all$label), function(lbl) {
  gr <- growth_all %>% filter(label == lbl) %>% select(zone_key, log_growth)
  zones_map %>%
    left_join(gr, by = "zone_key") %>%
    mutate(label = factor(lbl, levels = levels(growth_all$label)))
}))

ggplot() +
  geom_sf(data = zones_growth,
          aes(fill = log_growth),
          colour = NA) +
  geom_sf(data = province_outline,
          fill = NA, colour = "black", linewidth = 0.4) +
  geom_sf(data = cortina_pt,
          colour = "red", size = 3, shape = 16) +
  facet_wrap(~label, ncol = 2) +
  scale_fill_distiller(
    name      = "Log growth\n(≈ % change)",
    palette   = "RdYlGn",
    direction = 1,
    limits    = c(-scale_cap, scale_cap),
    oob       = scales::squish,
    na.value  = "grey92",
    labels    = function(x) paste0(ifelse(x > 0, "+", ""), round(x * 100, 1), "%")
  ) +
  labs(
    title    = "Log Price Growth: Before vs After Olympic Announcement",
    subtitle = "Abitazioni civili, Stato NORMALE | Black diamond = Cortina | Grey = no data",
    caption  = "Source: OMI / Agenzia delle Entrate"
  ) +
  annotation_north_arrow(
    location = "tl", which_north = "true",
    style = north_arrow_fancy_orienteering(
      fill = c("black", "white"), line_col = "black"
    ),
    height = unit(1.2, "cm"), width = unit(1.2, "cm")
  ) +
  coord_sf(datum = sf::st_crs(32632)) +
  theme_void(base_size = 11) +
  theme(
    plot.title          = element_text(size = 14, face = "bold"),
    plot.subtitle       = element_text(size = 9, colour = "grey40"),
    strip.text          = element_text(size = 11, face = "bold"),
    legend.position     = "right",
    legend.key.height   = unit(3, "cm"),
    panel.grid.major    = element_line(colour = "grey80", linewidth = 0.3),
    panel.grid.minor    = element_blank(),
    axis.text           = element_text(size = 7, colour = "grey50"),
    axis.text.x         = element_text(angle = 45, hjust = 1),
    panel.background    = element_rect(fill = "white", colour = NA),
    plot.background     = element_rect(fill = "white", colour = NA)
  )

dir.create(here::here("Figures"), showWarnings = FALSE)
ggsave(here::here("Figures", "map_prices_faceted.png"),
       width = 16, height = 9, dpi = 600, bg = "white")
cat("Saved to Figures/map_prices_faceted.png\n")
```

## 5c. Most Recent Period Prices (2025/S1)

```{r map-prices-latest, fig.height=9}
latest_prices <- panel %>%
  filter(year == 2025, sem == 1) %>%
  group_by(zone_key) %>%
  summarise(mean_price = mean(price_mid, na.rm = TRUE), .groups = "drop")

zones_map_prices <- zones_map %>%
  left_join(latest_prices, by = "zone_key")

ggplot() +
  geom_sf(data = zones_map_prices,
          aes(fill = mean_price),
          colour = "white", linewidth = 0.08) +
  geom_sf(data = province_outline,
          fill = NA, colour = "black", linewidth = 0.5) +
  geom_sf(data = cortina_pt,
          colour = "white", size = 4, shape = 18) +
  scale_fill_viridis_c(
    name     = "Price\n(€/m²)",
    option   = "inferno",
    na.value = "grey90"
  ) +
  labs(
    title    = "Residential Purchase Prices — Belluno Province (2025/S1)",
    subtitle = "Abitazioni civili, Stato NORMALE | White diamond = Cortina",
    caption  = "Source: OMI / Agenzia delle Entrate (2025)"
  ) +
  theme_void() +
  theme(
    plot.title    = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 10, colour = "grey40"),
    legend.position = "right"
  )
```

---

# 6. Price Trends by Distance Bin

## 6a. Level Prices Over Time

```{r price-trends}
trend_df <- panel %>%
  group_by(time, dist_bin) %>%
  summarise(mean_price = mean(price_mid, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(dist_bin))

ggplot(trend_df, aes(x = time, y = mean_price,
                     colour = dist_bin, group = dist_bin)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  geom_vline(xintercept = 2019.5, linetype = "dashed",
             colour = "firebrick", linewidth = 0.8) +
  annotate("text", x = 2019.65,
           y = max(trend_df$mean_price, na.rm = TRUE) * 0.97,
           label = "Olympic\nAnnouncement\nJun 2019",
           hjust = 0, size = 3, colour = "firebrick") +
  scale_colour_viridis_d(name = "Distance bin", option = "turbo") +
  scale_y_continuous(labels = label_comma(suffix = " €/m²")) +
  scale_x_continuous(breaks = seq(2014, 2025, 2)) +
  labs(
    title    = "Mean Residential Prices by Distance to Cortina",
    subtitle = "Province of Belluno — OMI Cod_Tip 20, Stato NORMALE",
    x = "Semester", y = "Mean price (€/m²)",
    caption = "Source: OMI / Agenzia delle Entrate (2014–2025)"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")
```

## 6b. Price Index (Base = 2019/S1)

```{r price-index}
base_df <- panel %>%
  filter(year == 2019, sem == 1) %>%
  group_by(zone_key) %>%
  summarise(base_log = mean(log(price_mid), na.rm = TRUE), .groups = "drop")

index_df <- panel %>%
  left_join(base_df, by = "zone_key") %>%
  filter(!is.na(base_log)) %>%
  mutate(log_diff = log(price_mid) - base_log) %>%
  group_by(time, dist_bin) %>%
  summarise(mean_log_diff = mean(log_diff, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(dist_bin))

ggplot(index_df, aes(x = time, y = mean_log_diff,
                     colour = dist_bin, group = dist_bin)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  geom_vline(xintercept = 2019.5, linetype = "dashed",
             colour = "firebrick", linewidth = 0.8) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.4) +
  annotate("text", x = 2019.65, y = max(index_df$mean_log_diff, na.rm = TRUE) * 0.9,
           label = "Olympic\nAnnouncement",
           hjust = 0, size = 3, colour = "firebrick") +
  scale_colour_viridis_d(name = "Distance bin", option = "turbo") +
  scale_x_continuous(breaks = seq(2014, 2025, 2)) +
  coord_cartesian(ylim = c(-0.05, 0.05)) +
  labs(
    title    = "Log Price Change by Distance to Cortina (Base = 2019/S1)",
    subtitle = "Log difference from base period — approx. % change",
    x = "Semester", y = "Log price difference",
    caption = "Source: OMI / Agenzia delle Entrate (2014–2025)"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")
```

## 6c. Price Index (Base = 2019/S1 = 100)

```{r price-index-level}
base_df2 <- panel %>%
  filter(year == 2019, sem == 1) %>%
  group_by(zone_key) %>%
  summarise(base_price = mean(price_mid, na.rm = TRUE), .groups = "drop")

index_df2 <- panel %>%
  left_join(base_df2, by = "zone_key") %>%
  filter(!is.na(base_price), base_price > 0) %>%
  mutate(price_index = price_mid / base_price * 100) %>%
  group_by(time, dist_bin) %>%
  summarise(mean_index = mean(price_index, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(dist_bin))

p_index <- ggplot(index_df2, aes(x = time, y = mean_index,
                      colour = dist_bin, group = dist_bin)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  geom_vline(xintercept = 2019.5, linetype = "dashed",
             colour = "firebrick", linewidth = 0.8) +
  geom_hline(yintercept = 100, colour = "grey50", linewidth = 0.4) +
  annotate("text", x = 2019.65, y = 118,
           label = "Olympic\nAnnouncement",
           hjust = 0, size = 3, colour = "firebrick") +
  scale_colour_viridis_d(name = "Distance bin", option = "turbo") +
  scale_x_continuous(breaks = seq(2014, 2025, 2)) +
  coord_cartesian(ylim = c(95, 105)) +
  labs(
    title    = "Price Index by Distance to Cortina (Base = 2019/S1 = 100)",
    subtitle = "Highlights divergence in trajectories post-announcement",
    x = "Semester", y = "Price index",
    caption = "Source: OMI / Agenzia delle Entrate (2014–2025)"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "bottom")

p_index

dir.create(here::here("Figures"), showWarnings = FALSE)
ggsave(here::here("Figures", "price_index.png"),
       plot = p_index, width = 10, height = 6, dpi = 300)
cat("Saved to Figures/price_index.png\n")
```

---

# 7. Distance-Price Gradient

## 7a. Pre vs Post Announcement

```{r gradient-plot}
gradient_df <- panel %>%
  mutate(period = if_else(post == 0,
                          "Pre-2019 (2014-2019/S1)",
                          "Post-2019 (2019/S2-2025/S2)")) %>%
  group_by(Comune_amm, Comune_descrizione, dist_cortina_km, period) %>%
  summarise(mean_price = mean(price_mid, na.rm = TRUE), .groups = "drop")

ggplot(gradient_df, aes(x = dist_cortina_km, y = mean_price, colour = period)) +
  geom_point(alpha = 0.45, size = 1.8) +
  geom_smooth(method = "loess", span = 0.5, se = TRUE, linewidth = 1.2) +
  scale_colour_manual(
    values = c("Pre-2019 (2014-2019/S1)"      = "steelblue",
               "Post-2019 (2019/S2-2025/S2)"  = "firebrick"),
    name = NULL
  ) +
  scale_y_continuous(labels = label_comma(suffix = " €/m²")) +
  labs(
    title    = "Distance-Price Gradient: Pre vs Post Olympic Announcement",
    subtitle = "Each dot = municipality average across all zones and semesters within period",
    x = "Euclidean distance to Cortina (km)",
    y = "Mean price (€/m²)",
    caption = "Source: OMI / Agenzia delle Entrate (2014–2025)"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank(), legend.position = "top")
```

## 7b. % Change (Post - Pre) vs Distance

```{r price-change-gradient}
change_df <- gradient_df %>%
  pivot_wider(
    id_cols     = c(Comune_amm, Comune_descrizione, dist_cortina_km),
    names_from  = period,
    values_from = mean_price
  ) %>%
  rename(
    pre  = `Pre-2019 (2014-2019/S1)`,
    post = `Post-2019 (2019/S2-2025/S2)`
  ) %>%
  mutate(price_change_pct = (post / pre - 1) * 100) %>%
  filter(!is.na(price_change_pct))

top_labels <- change_df %>%
  arrange(desc(abs(price_change_pct))) %>%
  slice_head(n = 8)

ggplot(change_df, aes(x = dist_cortina_km, y = price_change_pct)) +
  geom_hline(yintercept = 0, colour = "grey60", linewidth = 0.5) +
  geom_point(aes(size = abs(price_change_pct)),
             colour = "steelblue", alpha = 0.6) +
  geom_smooth(method = "loess", span = 0.5, se = TRUE,
              colour = "firebrick", linewidth = 1.2, fill = "firebrick") +
  ggrepel::geom_text_repel(
    data = top_labels,
    aes(label = Comune_descrizione),
    size = 2.8, colour = "grey30", max.overlaps = 15
  ) +
  scale_size_continuous(range = c(1, 7), guide = "none") +
  labs(
    title    = "% Price Change (Post vs Pre Announcement) by Distance to Cortina",
    subtitle = "Spatial decay of the Olympic spillover effect",
    x = "Euclidean distance to Cortina (km)",
    y = "% price change",
    caption = "Source: OMI / Agenzia delle Entrate (2014–2025)"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.minor = element_blank())
```

---

# 8. Econometric Baseline: TWFE

Two-way fixed effects (zone FE + semester FE).
Treatment: `dist_cortina_km * post` — the continuous DiD interaction.

```{r twfe}
reg_df <- panel %>%
  mutate(
    treat   = dist_cortina_km * post,
    zone_fe = zone_key,
    time_fe = paste0(year, "_", sem)
  ) %>%
  filter(!is.na(log_price))

# Model 1: linear distance x post
m1 <- feols(log_price ~ treat | zone_fe + time_fe,
            data = reg_df, cluster = ~zone_fe)

# Model 2: log distance x post
m2 <- feols(log_price ~ I(log1p(dist_cortina_km) * post) | zone_fe + time_fe,
            data = reg_df, cluster = ~zone_fe)

etable(m1, m2,
       headers  = c("Linear dist x post", "Log dist x post"),
       title    = "TWFE Baseline — Dependent Variable: log(price)",
       digits   = 4)
```

---

# 9. Event Study (Dynamic DiD)

Interact distance with each semester relative to 2019/S1 to verify
pre-announcement parallel trends.

```{r event-study}
reg_df <- reg_df %>%
  mutate(rel_time = (year - 2019) * 2 + (sem - 1))

m_es <- feols(
  log_price ~ i(rel_time, dist_cortina_km, ref = 0) | zone_fe + time_fe,
  data    = reg_df,
  cluster = ~zone_fe
)

iplot(m_es,
      main = "Event Study: Distance x Semester Coefficients",
      xlab = "Semesters relative to Olympic Announcement (0 = 2019/S1)",
      ylab = "Coefficient on log price",
      sub  = "95% CI — clustered SE at zone level")
abline(v = 0.5, col = "firebrick", lty = 2)
```

---

# 10. Summary Statistics Table

```{r summary-table}
panel %>%
  group_by(dist_bin) %>%
  summarise(
    Observations  = n(),
    Zones         = n_distinct(zone_key),
    Municipalities = n_distinct(Comune_amm),
    `Mean price`  = round(mean(price_mid, na.rm = TRUE)),
    `Median price` = round(median(price_mid, na.rm = TRUE)),
    `SD price`    = round(sd(price_mid, na.rm = TRUE)),
    `Mean dist km` = round(mean(dist_cortina_km, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  knitr::kable(
    caption = "Summary Statistics by Distance Bin to Cortina d'Ampezzo",
    format.args = list(big.mark = ",")
  )
```

---

# Session Info

```{r session-info}
sessionInfo()
```
